Fixed-Parameter Tractable Sampling for RNA Design with Multiple Target Structures

### 2018-10-08 RNARedPrint.tex
sync

\section*{Results}\label{sec:results}
Our Boltzmann sampling strategy (Algorithms
\ref{alg:pf} and \ref{alg:sampling}) was implemented, using a combination of {\tt C++} and {\tt Python}, in a software called \ourprog{}, available at:

@@

We implemented the core algorithms in \texttt{C++}, resulting in the tool
\ourprog{}, which takes a set of target structures, as well as
weights for each energy feature and \GCb\%, and generates a sample set of sequences compatible with the structures in the corresponding
Boltzmann distribution. \ourprog{} currently supports the base pair
energy model and the stacking energy model.
Moreover, we provide a {\tt Python} wrapper script that targets prescribed energies using multi-dimensional Boltzmann sampling.
For a given set of secondary structures, together with prescribed target energies
and target \GCb\%, this script generates a series of sequences that satisfy the target values for the energy and
\GCb\% features within configurable tolerances.
Notably, these target energies are actual free energies in the realistic
Turner energy model, and are targeted by efficiently sampling in the
base pair energy model, and filtering sequences based on the {\tt RNAeval} tool from the Vienna package~\cite{Lorenz2011}.

@@

We considered a benchmark of six sets of target structures described in \citep{Taneda2015}: {\sf 2str}, {\sf 3str}, and {\sf 4str} consist of non-pseudoknotted structures, while {\sf PK60}, {\sf PK80}, and {\sf LE80} contain pseudoknotted structures.
%%For each instance of the \Software{Modena}
%benchmark~\citep{Taneda2015}, we generated $1\,000$ seed sequences at these energies.
%Our procedure is tailored to produce sequences with similar Turner
%energy that favor the stability of the target structures having
%moderate \GCb\%; all these properties are desirable objectives
%for RNA design.
Using \ourprog{}, we generated at least 1\,000 seeds for
each instance of the benchmark. We used target energies that were
computed by averaging the energy over samples at higher weights
$e^{\beta}$, and set the targeted \GCb\% to 60\%.  We compared the
\Obj{} value of these sequences against that of seed sequences,
uniformly sampled using \RNAblueprint{}.  Moreover, for both sets, we
used an adaptive greedy walk~\citep{Hammer2017} to \Def{minimize} the
\Obj{} function. At each step, the local search resamples (uniformly
at random) the positions of a randomly selected component in the base
pair dependency graph, accepting the modification only if it results
in a gain. We performed 500 greedy descent steps in the case of
pseudoknot-free datasets {\sf 2str}, {\sf 3str}, and {\sf 4str}; and
200 steps for the pseudoknotted ones {\sf PK60}, {\sf PK80}, and {\sf
  LE80}.

>>>>

\section*{Results}\label{sec:results}

We implemented the core algorithms in \texttt{C++}, resulting in the tool
\ourprog{}, available at:

@@

\ourprog{}  takes a set of target structures, as well as
weights for each energy feature and \GCb\%, and generates a sample set of sequences compatible with the structures in the corresponding
Boltzmann distribution. \ourprog{} currently supports the base pair
energy model and the stacking energy model.

Moreover, we provide a {\tt Python} wrapper script that targets prescribed energies using multi-dimensional Boltzmann sampling.
For a given set of secondary structures, together with prescribed target energies
and target \GCb\%, this script generates a series of sequences that satisfy the target values for the energy and
\GCb\% features within configurable tolerances.
Notably, these target energies are actual free energies in the realistic
Turner energy model, and are targeted by efficiently sampling in the
base pair energy model, and filtering sequences based on the {\tt RNAeval} tool from the Vienna package~\cite{Lorenz2011}.

@@

We considered a benchmark of six sets of target structures described in \citep{Taneda2015}: {\sf 2str}, {\sf 3str}, and {\sf 4str} consist of non-pseudoknotted structures, while {\sf PK60}, {\sf PK80}, and {\sf LE80} contain pseudoknotted structures.
%%For each instance of the \Software{Modena}
%benchmark~\citep{Taneda2015}, we generated $1\,000$ seed sequences at these energies.
%Our procedure is tailored to produce sequences with similar Turner
%energy that favor the stability of the target structures having
%moderate \GCb\%; all these properties are desirable objectives
%for RNA design.
Using \ourprog{}, for
each instance of the benchmark, we generated at least 1\,000 seed sequences with similar
energies for all target structures. For this purpose, we determine good common target energies that
can be succesfully targeted for all single target structures simultaneously.
Generally, we targeted 60\% \GCb\%.  We compared the
\Obj{} value of these sequences against that of seed sequences,
uniformly sampled using \RNAblueprint{}.  Moreover, for both sets, we
used an adaptive greedy walk~\citep{Hammer2017} to \Def{minimize} the
\Obj{} function. At each step, the local search resamples (uniformly
at random) the positions of a randomly selected component in the base
pair dependency graph, accepting the modification only if it results
in a gain. We performed 500 greedy descent steps in the case of
pseudoknot-free datasets {\sf 2str}, {\sf 3str}, and {\sf 4str}; and
200 steps for the pseudoknotted ones {\sf PK60}, {\sf PK80}, and {\sf
  LE80}.
